---
title: "Homework 6"
subtitle: "Autumn 2022"
author: "KEY"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * * 

### Instructions

- This homework is due in Gradescope on Wednesday Nov 16 by midnight PST.

- Please answer the following questions in the order in which they are posed. 

- Don't forget to knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. 


* * *


1. The number of chocolate chips in a certain type of cookie has a Poisson distribution. We want the probability that a randomly chosen cookie has at least 2 chocolate chips to be greater than 0.99. 

a. Make a line plot of this probability as a function of $\lambda$. 
Let $X$ denote the number of chocolate chips in a randomly selected cookie. We are given
$$X \sim Pois(\lambda).$$
This means
$$f(x) = \lambda^{x} e^{-\lambda x}/x!, \ \ x = 0,1,2\dots$$
Therefore
\begin{align*}
P(X \geq 2) &= 1 - P(X< 2) = 1 - P(X = 0) - P(X = 1)\\
&= 1 - e^{-\lambda} - \lambda e^{-\lambda}.
\end{align*}
In the following code chunk, we will make a plot of $P(X \geq 2)$ versus $\lambda$.

```{r}
ggplot() + 
  geom_function( 
    fun = function(x) {1 - exp(-x) - x*exp(-x)},
    xlim=c(0,10)  #by 10 the function is pretty stable
    )+
  labs(
    x = expression(lambda),
    y = "P(X >=2)",
    title = "Probability of at least 2 chocolate chips "
  )


```

b. Find the smallest value of the parameter $\lambda$ of this distribution that ensures this probability.

(Modify the \texttt{uniroot} code below. You can pick 0 for a lower bound for the mean and 10 for the upper bound )


```{r  message= F, warning = F}
#The function uniroot searches the interval from lower to upper
#for a root of the function. Type ?uniroot for more information.

#to find x so that x(x^2-1)+0.5=0 in the range -2, 2
##use the following code

uniroot(f = function(x){ 1 - exp(-x) - x*exp(-x)} - 0.99, lower=0, upper=10)  

```

The smallest value of $\lambda$ is 6.638. This is the root of the equation
$$1 - e^{-\lambda} - \lambda e^{-\lambda} = 0.99$$.
For any $\lambda$ that is larger than 6.638, this probability will be exceeded. 

2. (Symmetry) Suppose that $f(x)$ is the PDF corresponding to a continuous random variable which is symmetrically distributed about 0. That is, for any $a > 0$
   $$f(a) = f(-a).$$
	Show that 
	$$P(-a \leq X \leq a) = 2 F(a) - 1$$ where $F(x)$ is the CDF corresponding to  $f(x)$. 

    \emph{Hint:} 
     $$P(-a \leq X \leq a) = P(X \leq a) - P(X < -a).$$ 
     Write
     $$P(X < -a ) = \int\limits_{-\infty}^{-a} f(x) dx.$$ 
     Make a change of variable $u = -x$ and proceed from there.

We begin as suggested in the hint:
\begin{align*}
P(-a \leq X \leq a) &= P(X \leq a) - P(X < a)\\
&= F(a) - P(X<a).
\end{align*}
We will now show that $P(X < a) = 1 - F(a)$. By definition:
$$P(X < a) = \int\limits_{-\infty}^{a} f(x)\:dx.$$
Let's make the change of variable
$$u = -x \Rightarrow du = -dx$$
in the integral above.
This leads to
\begin{align*}
P(X < a) 
&= \int\limits_{\infty}^{-a} f(-u) -du,\\
&= \int\limits_{a}^{\infty} f(-u) du, \ \ \text{flip limits} \\
&= \int\limits_{a}^{\infty} f(u) du \ \ \text{f is symmetric}.\\
&= P(X > a) \\
&= 1 - P(X \leq a) = 1 - F(a).
\end{align*}
  
4. (Flooding river) A river floods every year. Suppose that the low-water mark is set at 1 and the high-water mark $X$ has CDF
\begin{align*}
	F_X(x) &= P(X \leq x) \\
	       &= \left\{ \begin{array}{cc} 
		               0 & x < 1 \\
		               1 - 1/x^2 & 1 \leq x < \infty. \end{array} \right.
\end{align*}

a.  Find a PDF, $f(x)$ for $X$.

By the fundamental rule of calculus
$$f(x) = \frac{d}{dx} F(x).$$
Therefore, for $x > 1$ we have
\begin{align*}
f(x) &= \frac{d}{dx} \left(1- \frac{1}{x^2} \right), \\
&= \frac{d}{dx} -x^{-2},\\
&= 2 x^{-3}.
\end{align*}
Note that it is assumed that for $x \leq 1$, the PDF is 0.

b. If the low-water mark is set to 0 and we use a unit of measurement that is $\frac{1}{10}$ of that given previously, the high-water mark becomes 
     $$Y=10\:(X-1).$$ 

     
   Find $P(Y \leq 1)$. 
   
\begin{align*}
P(Y \leq 1) &= P(10\:(X - 1) \leq 1), \\
&= P(X \leq \frac{1}{10} + 1 ),\\
&= F(\frac{1}{10} + 1),\\
&= 1 - 1/(\frac{1}{10} + 1)^2,  \ \ F(x) = 1-1/x^2 \ x \geq 1.\\
&= `r 1 - 1/(11/10)^2`.
\end{align*}

4. The following ordered set of 27 P-values (from Kaati et al., Eur. J. Hum Genetics 2007) were the result of testing many independent subgroups of a sample. 

```{r pvalues}
pvalue_df <- data.frame( 
  pval = c(0.01, 0.01, 0.02, 0.04, 0.04, 0.05, 0.07, 0.07, 0.10, 0.19, 0.24, 0.27, 0.34,
0.37, 0.44, 0.50, 0.53, 0.54, 0.55, 0.61, 0.70, 0.77, 0.80, 0.80, 0.82, 0.94, 0.99)
)
```

It is hypothesized that the P-values all come from a
uniform distribution on $[0,1)$. Fit a uniform distribution to these data. How well does it fit (just visually)?

The following code creates the histogram on the density scale with the uniform distribution superimposed. Visually, the uniform is not a good fit.

```{r}

ggplot(data = pvalue_df) +
  geom_histogram( 
       mapping = aes(x = pval, y = ..density..),
       bins = 6, #sturges: num of bins = log2(27) + 1 approx 6 
       color = "purple",
       fill = "gold",
       breaks = seq(0,1,1/6)) +
  geom_function(fun = dunif,
                args=list(min = 0, max = 1),
                xlim=c(0,1) ) +
  labs(
    title = "Pvalues with uniform distribution superimposed"
  )
```