---
title: "Modeling fumbles"
author: " "
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)               #for functions
library(fastR2)                  #for data
```


1. Load the data `Fumbles` into the Environment by typing `data(fumbles)` in the Console. Take a look at it in the spreadsheet viewer. 

2. Create the frequency distribution of the number of fumbles in `week2`   

3. Using `summarize` print the mean and standard deviation of the number of fumbles made in `week2`. 

4. Create a new dataframe called `pois_fit` that contains the probabilities for the Poisson distribution with $\lambda$ equal to the mean from 3. 

5. Create a histogram of the number of fumbles for `week2` on a density scale. Superimpose the Poisson probabilities on the histogram. Is it a good fit?

6. (If we have time) Repeat parts 1 through 5 for a new variable - `total` - the total number of fumbles across all three weeks. 

```{r create-total}

Fumbles <- Fumbles %>% 
             mutate(total = week1 + week2 + week3)
```