---
title: "Problem Session 5 KEY"
subtitle:  "Negative binomial"
graphics: yes
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. Focus on answering in complete sentences. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 

1. (Mean and variance) With a Nevada Roulette, there are 38 numbers the wheel could land on when spun: 0, 00, 1, 2, $\dots$, 36. The single number bet has a payout of 35 to 1: which means you'll win \$35 for every dollar you bet . Suppose you bet \$5 on your favorite number. Let $X$ denote your \emph{net gain} in one play. The PMF of $X$ is shown in the table below.

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|} \hline
$x$ & -\$5 & \$170 \\ \hline 
$f(x)$ & 37/38 & 1/38  \\ \hline
\end{tabular}
\end{table}

   Calculate your expected net gain? By how much will your net gain typically vary from expected? Calculate the standard deviation of the net gain.
   
   We may solve for $E(X)$ using the definition:
   
   $E(X) = \sum_x xf(x) = -5(\frac{37}{38}) + 170(\frac{1}{38}) = \$`r -5*37/38 + 170/38`$
   
   We know $Var(X) = E(X^2) - E(X)^2$. We know $E(X)$, so we can now find $E(X^2)$
   
   $E(X^2) = \sum_x x^2f(x) = (-5)^2(\frac{37}{38}) + 170^2(\frac{1}{38}) = `r 25*37/38 + 170^2/38`$
   
   Thus we have that $Var(X) = `r 25*37/38 + 170^2/38 - (-5*37/38 + 170/38)^2`$. We thus have that $SD(X) = \$`r sqrt(25*37/38 + 170^2/38 - (-5*37/38 + 170/38)^2)`$.
   
```{r}
E_X <- -5*37/38 + 170/38
sd_X <- sqrt(25*37/38 + 170^2/38 - (-5*37/38 + 170/38)^2)
```
   
   
   The SD tells us the typical deviation in earnings from the average earnings ($E(X)$). Typically the values will be in the interval $[E(X)-SD(X),E(X)-SD(X)] = [`r E_X - sd_X`,`r E_X+sd_X`]$

2. (Chebychev) Suppose $X$ is a random variable with mean and variance both equal to 20. What can be said about $P( 0 <  X < 40)$?

  We have that:
  \begin{align*}
  P( 0 <  X < 40) & \\
  &= P(-20 < X-20 < 20) \\
  &= P(\frac{-20}{\sqrt{20}} < \frac{X-20}{\sqrt{20}} < \frac{20}{\sqrt{20}}  ) \\
  &= P(-\sqrt{20} < \frac{X-20}{\sqrt{20}} < \sqrt{20}) \\ 
  &= P(\frac{|X-20|}{\sqrt{20}} < \sqrt{20}) \\
  &= 1 - P(\frac{|X-20|}{\sqrt{20}} \geq \sqrt{20}) \\
  &= 1 - P(|X-20| \geq \underbrace{\sqrt{20}}_{\sigma} \times \underbrace{\sqrt{20}}_k) \\
  &\geq  1-\frac{1}{\underbrace{\sqrt{20}^2}}_{k^2} \\ 
  &= `r 1-(1/20)`
  \end{align*}

3. (Negative binomial) Suppose we wish to generate $X \sim Binom(10, \frac{1}{2})$ subject to the constraint $X \leq 3$. Say we use the following naive algorithm to accomplish this task:

  - Generate $x$ from a $Binom(10, \frac{1}{2})$
  
  - Accept the value $x$ if $x <= 3$. Otherwise reject it.
  
a. Calculate the acceptance probability.

  We accept $x$ if $x\leq3$. Thus the acceptance probability is $P(X\leq3)$. We can find this using `pbinom`, and we have that $P(X\leq3) = `r pbinom(3,10,.5)`$

b. Suppose we actually want 25 acceptable x values. Let $Y$ denote the number of rejections before we find our 25 acceptable values. What is the distribution of $Y$? 

$Y$ is a negative binomial distribution, so $Y\sim NBinom(s=25,\pi = `r pbinom(3,10,.5)`)$

c. Make the probability histogram of $Y$. Just eyeballing, how many rejections are most likely before we get our acceptable 25?

```{r}
data.f <- tibble(
  x = 0:150,
  f = dnbinom(x, size = 25, prob = pbinom(3,10,0.5) )
)

ggplot(data.f, aes(x=x,y=f))+geom_col(width=1, alpha=0.5)
```

The histogram peaks around X=115, indicating that is the most likely number of rejections.

d. Generate $y+25$ binomials where $y$ is your answer for the most likely number of rejections from part c. How many do you reject?

```{r}
set.seed(1414)

y<-rbinom(n=145,size=10,prob=0.5)
sum(y >3)
```


4. (Geometric series) Two golfers are playing sudden death to decide a tournament. The first one wins a hole with probability p, the second one wins with probability q, and holes are tied with probability r. Holes are independent, and the game stops the first time someone wins a hole. What is the probability that the first player wins?

Player A wins the game if they win the game first. 

Let X be the number of games it takes for player A to win. Then X can be 1, 2 (tie, win), 3(tie, tie, win), ... .

P(Player A Wins) = $P(X=1\cup X=2 \cup \dots) = \sum_{i=1}^{\infty}P(X=i)$

The probability player A wins in $X=x$ games is $r^{x-1}p$, where $x=1,2,\dots$.

Thus P(Player A Wins) = $\sum_{x=1}^{\infty} r^{x-1}p$

So we have that:
\begin{align*}
P(A) &= \\ 
&= \sum_{x=1}^{\infty} r^{x-1}p \\
&= \sum_{x=1}^{\infty} r^{x-1}p \\
&= (p + pr + pr^2 + \dots) \\
&= p\frac{1}{1-r}
\end{align*}

Here we use the identity:

$\sum_{k=0}^{\infty} ar^k =  a + ar + ar^2 \dots = \frac{a}{1-r}$ for $r<1$. Here we have a fraction with the numerator as the common term of the series, and the denominator as $1-r$ where r is the geometric rate. 





