---
title: "Problem Session 4 Key"
subtitle:  "Binomial Distribution & Mean of a Random Variable"
graphics: yes
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. Focus on answering in complete sentences. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 

1. A box contains n chips numbered 1 through n.
Assume that the probability of choosing chip i is equal
to $k \cdot i,i = 1, 2,\dots, n$.

a. What is $k$?  (Hint: the sum of the first $n$ integers is $n(n+1)/2$)

We know that for any probability distribution, f(x), for a discrete RV X, we must have: $\sum_x f(x) = 1$

We may use this property to solve for k.

\begin{align*}
\sum_x f(x) &= 1 \\
\sum_{i=1}^n ki &= 1 \\
k\sum_{i=1}^n i &= 1 \\
k\frac{n(n+1)}{2} &= 1 \\
k &= \frac{2}{n(n+1)}
\end{align*}

b. Let $X$ denote the number on a randomly selected chip from the box. Write its PMF.

We have that $f(x) = P(X=i) = ki = \frac{2i}{n(n+1)}$

c. Give an expression for $E\left[X \right]$. Put a box around your final answer. (Hint: the sum of the squares of the first $n$ integers is $n(n+1)(2n+1)/6$)

\begin{align*}
E[X] &= \sum_x f(x) \\
&= \sum_{x=1}^n \frac{2x^2}{n(n+1)} \\
&= \frac{2}{n(n+1)}\sum_{x=1}^n x^2 \\
&= \frac{2}{n(n+1)}\frac{n(n+1)(2n+1)}{6} \\
&= \frac{1}{3(2n+1)}
\end{align*}

So $E[X] = \frac{1}{3(2n+1)}$

2. A manufacturer of computer memory chips claims they have a 4% defective rate. A factory which assembles laptops orders 20 chips from them. Let the random variable $X$ denote the number of defective chips in the order. 

a. What is the distribution of $X$? State the name of the distribution and also the values of its parameters. Be sure to state any assumptions you are making when deciding on the distribution.

We have 4 conditions for a binomial experiment.

(i) Our trial is randomly selecting one chip from a manufacturers lot and checking to see if it is defective. Thus n=20, the number of trials, is specified.

(ii) For each trial we either have a defective chip or not, thus each trial has 2 outcomes, where we define finding a defective chip as a success.

(iii) We see each chip has a 0.04 chance of failing, and thus each trial has the same probability of success. 

In this framework we select 20 chips without replacement. We would expect trials would change probability from each trial to each trial, but as we saw in lecture, the probability of success on any given trial (for example trial 2) is the same for each trial, that is 0.04 because we average over the previous trials. Refer to example 6.1 from lecture for further information.   

(iv) The last assumption deals with the question of are the trials independent. Since this is sampling without replacement, the answer is no. BUT, if we assume N (the total number of chips) is sufficiently large, then we get approximate independence. To see why lets calculate the probability that the second draw is defective given the first is defective. Since 0.04*N chips are defective, the probability is $\frac{0.04N - 1}{N-1} = 0.04 + \frac{.04-1}{N-1}$. If N is large this probability will be very close to 0.04 (in fact as N approaches infinity this probability converges exactly to 0.04).

Thus X is a binomial random variable with the following parameters:

$$X \sim Bin(n=20,\pi=0.04)$$

b. How many defective chips should the factory *expect* to receive?

We know $E(X)=n\pi$ for a binomial random variable X

Thus we have that:

$$E(X) = n\pi = 20*(0.04) = .8$$

c. Calculate the probability that they will receive more defective chips than expected?

For this we wish to calculate $P(X>.8)$
\begin{align*}
P(X>0.8) &= P(X\geq 1) \\ 
&= 1 - P(x<1) \\
&= 1 - P(X=0) \\
&= 1 - (0.96)^{20} \\
&= `r 1 - (0.96)^(20)`
\end{align*}

We may also utilize the cdf function of the binomial in R, pbinom to solve this. 

```{r}
pbinom(0.8,size=20,p=.04,lower.tail = F)
```

3. Weird Wally offers you the following choice. You may have 1/3.5 dollars or you may roll a fair die and he will give you 1/X dollars where $X$ is the value of the roll. Which is the better deal? Calculate $E\left[ \frac{1}{X} \right]$ to decide.

We can find $E\left[\frac{1}{X}\right]$ as follows by using the Law of the Unconscious Statistician/Probabilist:

\begin{align*}
E\left [\frac{1}{X}\right ] &= \sum_x \frac{1}{x}f(x) \\ 
&= \sum_{x=1}^6 \frac{1}{x} \frac{1}{6} \\ 
&= \frac{1}{6}\sum_{x=1}^6\frac{1}{x}
\end{align*}

We may calculate this using R.

```{r}
inv_x <- 1/(1:6)
E_1_x <- sum(inv_x)/6
```

Using R we see that $E\left[ \frac{1}{X} \right ] = `r E_1_x`$ versus Wally's offer of `r 1/3.5`. Thus rolling the die has a better expected value, and is this the better deal. 
