---
title: "Problem Session 3"
subtitle:  "Discrete Distributions: PMF and CDF"
graphics: yes
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. Focus on answering in complete sentences. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 
	 
1. Suppose you toss a fair coin 3 times. The sample space for this experiment is shown below:
$$S = \{ HHH, HHT, HTH, HTT, TTT,TTH, THT, THH \}.$$

a. Write the PMF of $X$ defined as the difference between number of heads and number of tails.

\begin{tabular}{|c|c|c|c|c|} \hline
$x$ & -3 & -1 & 1 & 3 \\ \hline
$f(x)$ & 1/8 & 3/8 & 3/8 & 1/8 \\ \hline
\end{tabular}

b. Give the CDF of $X$.

\begin{align*} 
F(x) &= P(X\leq x) \\ 
&= \left \{ 
\begin{array}{cc}
                0 &\text{ if }  x < -3 \\ \\
      \frac{1}{8} &\text{ if } -3 \leq x < -1 \\ \\
      \frac{4}{8} &\text{ if } -1 \leq x < 1 \\ \\
      \frac{7}{8} &\text{ if } 1 \leq x < 3 \\ \\
                1 &\text{ if } x \geq 3
\end{array} \right.
\end{align*}

c. Draw the probability histogram of $X$ in R.

```{r}
library(tidyverse)

ht_df <- data.frame(
  x = c(-3,-1,1,3),
  probs = c(1/8,3/8,3/8,1/8))

ggplot(data=ht_df, mapping = aes(x=x,y=probs))+
  geom_col(width=1) + 
  labs(title = "Probability Histogram of X",
       y = "Probability of X",
       x = "X=x") #+
  #scale_x_continuous(breaks = c(-3,-1,1,3))
```

2. Suppose $X$ is a discrete uniform random variable on the integers $-n, -n + 1, -n +2, \dots, 0, 1, 2,\dots, n$, that is, 
$$f(x) = P(X = x) = c, \ \ \ x = -n, -n+1, -n+2, \dots, 1,2,\dots,n.$$

a. What is $c$?

We know c must be positive, and by definition of PMF's we must have that $\sum_{x}f(x) = \sum_{x=-n}^{x=n}c = 1$. We see there are $n$ elements in $1, \dots, n$, and likewise $n$ elements in $-n,\dots,-1$, and thus adding in 0, there are $2n+1$ elements in this space. Thus we have that $c=\frac{1}{2n+1}$

b. What is the PMF of  $$Y = -X? $$

The possible values for Y are: $y = -n, -(n-1), \dots, 0, 1, \dots, n$. This is the same range as the random variable X. 

we have that $P(Y=-n) = P(X=n) = c = \frac{1}{2n+1}$. This will extend to all values in the range of Y. 

Thus the PMF of $Y$ will simply be the PMF of $X$ since the space is symmetric around 0 and has uniform probabilities. So we have:
$$f(y) = P(Y = y) = c = \frac{1}{2n+1}$$  
	  
